Fourmile

### 2022-10-12 fourmile_dataprocess.Rmd
reorganized files to uncomment code

---
title: "Fourmile"
author: "Megan Sears"
date: "7/27/2022"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

@@

#save.image(file = 'spatial.RData')
load('C:/Users/sears/Documents/Repos/fourmile/data/spatial.RData')

@@

```{r, echo = F}
# niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>% 
#   filter(!row_number() %in% c(1:10)) %>%
#    rename(date = 1,
#           prism_mm = 2) %>%
#    mutate(date = ymd(date),
#           prism_mm = as.numeric(prism_mm))
# 
# niwot_snotel <- prec_update %>%
#   dplyr::select(c(date, Niwot_Snotel)) %>%
#   rename(snotel_mm = Niwot_Snotel) %>%
#   mutate(date = dmy(date))
# 
# niwot_compare <- full_join(niwot_prism, niwot_snotel, by = 'date')
# 
# niwot_compare <- niwot_compare %>%
#   mutate(diff_mm = snotel_mm - prism_mm) %>%
#   na.omit()
# 
# #pbias(niwot_compare$prism_mm, niwot_compare$snotel_mm)
# 
# scatter <- ggplot(niwot_compare, aes(x=snotel_mm, y=prism_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism vs. Niwot SNOTEL precip')
#   
# ggplotly(scatter)
# 
# rmse(niwot_compare$prism_mm, niwot_compare$snotel_mm)
# bias(niwot_compare$snotel_mm, niwot_compare$prism_mm)
# summary(niwot_compare$diff_mm)
# #filter to look at sept - june
# niwot_winter_comp <- niwot_compare %>%
#   mutate(month = month(date)) %>%
#   filter(!month %in% c(6:8))
# 
# summary(niwot_winter_comp)
```

@@

```{r, echo = F}
# niwot_prismshift <- niwot_compare %>%
#   mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
#   mutate(month = month(date)) %>%
#   #filter(!month %in% c(6:8)) %>%
#   mutate(diff_mm = snotel_mm - prism_mm) %>%
#   na.omit()
# 
# #NSE(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# #pbias(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# 
# scatter_shiftday <- ggplot(niwot_prismshift, aes(x=snotel_mm, y=prism_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift 1 day versus Niwot SNOTEL precip')
#   
# ggplotly(scatter_shiftday)
# 
# rmse(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# bias(niwot_prismshift$snotel_mm, niwot_prismshift$prism_mm)
# summary(niwot_prismshift$diff_mm)
```

@@

```{r, echo = F}
# niwot_prismfrac <- niwot_compare %>%
#   mutate(prism_most_mm = prism_mm * 0.974,
#          prism_little_mm = prism_mm * 0.026) %>%
#   mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
#   mutate(prism_new_mm = prism_most_mm + prism_little_mm,
#          diff_prismfrac = snotel_mm - prism_new_mm) %>%
#   na.omit()
# 
# scatter_frac <- ggplot(niwot_prismfrac, aes(x=snotel_mm, y=prism_new_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift by 97.4% and Niwot SNOTEL precip')
#   
# ggplotly(scatter_frac)
# 
# rmse(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
# bias(niwot_prismfrac$snotel_mm, niwot_prismfrac$prism_new_mm)
# summary(niwot_prismfrac$diff_prismfrac)
#NSE(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
#pbias(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)

@@

```{r, echo = F}
# srad <- solr %>%
#   dplyr::select(-1) %>%
#   filter(!row_number() %in% c(1:17)) %>%
#   replace(.==-9999, NA) %>%
#   rename(date = 1,
#          prism_w = 2,
#          prism_m = 3,
#          prism_e = 4,
#          gordon_s = 5,
#          gordon_n = 6,
#          greenlake = 7) %>%
#   mutate(date = dmy(date))
# 
# niwot_srad <- Niwo %>%
#   mutate(date = ymd(date)) %>%
#   rename(niwot_flux_mjm2 = 3) %>%
#   dplyr::select(-c(watts_per_m2)) %>%
#   replace(. == -9999, NA)
# 
# ## original  
# srad_compare <- full_join(srad, niwot_srad, by = 'date') %>%
#   dplyr::select(c(date, prism_w, niwot_flux_mjm2)) %>%
#   mutate(diff = niwot_flux_mjm2 - prism_w) %>%
#   na.omit()
# 
# scatter_srad_orig <- ggplot(srad_compare, aes(x=niwot_flux_mjm2, y=prism_w)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism versus Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_orig)
# 
# rmse(srad_compare$prism_w, srad_compare$niwot_flux_mjm2)
# bias(srad_compare$niwot_flux_mjm2, srad_compare$prism_w)
# summary(srad_compare$diff)
```

@@

```{r, echo = F}
## 1 day shift
# srad_shift1day <- srad_compare %>%
#   mutate_at(c("prism_w"),list(lead), n = 1) %>%
#   na.omit() %>%
#   mutate(diff = niwot_flux_mjm2 - prism_w)
# 
# scatter_srad_1day <- ggplot(srad_shift1day, aes(x=niwot_flux_mjm2, y=prism_w)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift 1 day vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_1day)
# #scatter_srad_1day
# 
# rmse(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
# bias(srad_shift1day$niwot_flux_mjm2, srad_shift1day$prism_w)
# summary(srad_shift1day$diff)
#pbias(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
```

@@

```{r, echo = F}
## shift by percentage
# srad_fracshift <- srad_compare %>%
#     mutate(prism_most = prism_w * 0.3,
#          prism_little = prism_w * 0.7) %>%
#   mutate_at(c('prism_most'), list(lead), n = 1) %>%
#   mutate(prism_new = prism_most + prism_little,
#          diff_prismfrac = niwot_flux_mjm2 - prism_new) %>%
#   na.omit()
#   
# scatter_srad_frac <- ggplot(srad_fracshift, aes(x=niwot_flux_mjm2, y=prism_new)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift by 30% vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_frac)
# #scatter_srad_frac
# 
# rmse(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
# bias(srad_fracshift$niwot_flux_mjm2, srad_fracshift$prism_new)
# summary(srad_fracshift$diff_prismfrac)
#pbias(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
```

@@

```{r, echo = F}
## bring in daymet
## Daily total radiation (MJ/m2/day) can be calculated as follows: ((srad (W/m2) * dayl (s/day)) / l,000,000)
# niwot_daymet <- daym %>%
#   filter(!row_number() %in% c(1:7)) %>%
#   rename(year = 1,
#          doy = 2,
#          srad_wm2 = 3) %>%
#   mutate(date = as.Date(doy),
#          month = month(date),
#          day = day(date),
#          date = make_date(year, month, day),
#          daymet_mjm2 = (srad_wm2 * 86400) / 1000000) %>%
#   dplyr::select(-c(year, doy, month, day))
#   
# niwot_daymet_comp <- full_join(niwot_srad, niwot_daymet, by = 'date') %>%
#   dplyr::select(c(date, niwot_flux_mjm2, daymet_mjm2)) %>%
#   na.omit() %>%
#   mutate(diff = niwot_flux_mjm2 - daymet_mjm2)
#   
# scatter_srad_daymet <- ggplot(niwot_daymet_comp, aes(x=niwot_flux_mjm2, y=daymet_mjm2)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Daymet vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_daymet)
# #scatter_srad_frac
# 
# #summary(srad_fracshift)
# rmse(niwot_daymet_comp$daymet_mjm2, niwot_daymet_comp$niwot_flux_mjm2)
# bias(niwot_daymet_comp$niwot_flux_mjm2, niwot_daymet_comp$daymet_mjm2)
# summary(niwot_daymet_comp$diff)
```

@@

```{r, echo=F}
#bring in srad and niwot_srad
# srad_gapfil <- full_join(srad, niwot_srad) %>%
#   filter(date > '2009-12-31') %>%
#   dplyr::select(-c(prism_w, prism_m, prism_e))
# 
# #bat <- read_xlsx('bat_srad.xlsx')
# #save.image(file = 'bat.RData')
# load('C:/Users/sears/Documents/Repos/fourmile/data/bat.RData')
# 
# bet_day <- bat %>%
#   mutate(date = as.Date(datetime)) %>%
#   group_by(date) %>%
#   summarize(sw_in_avg_wm2= mean(in_swrad_wm2)) %>%
#   filter(date > '2009-12-31') %>%
#   mutate(bet_sw_in_mjm2 = (sw_in_avg_wm2 * 86400) / 1000000) %>%
#   dplyr::select(-c(sw_in_avg_wm2))
# 
# srad_gapfil_all <- full_join(srad_gapfil, bet_day, by='date') %>%
#   filter(date < '2020-12-31') %>%
#   dplyr::select(-c(bet_sw_in_mjm2.x)) %>%
#   rename(bet_sw = bet_sw_in_mjm2.y) %>%
#   mutate(doyh = yday(date))
# 
# summary(srad_gapfil_all)
# 
# srad_flip_nofill <- srad_gapfil %>%
#   pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')
# 
# srad_nofill <- ggplot(srad_flip_nofill, aes(x=date, y=sw_in_mjm2, color=site)) +
#   geom_line() +
#   ggtitle('Solar rad, no fill')
# 
# ggplotly(srad_nofill)
# 
# new_srad <- missRanger(srad_gapfil_all, num.trees=100000)
# summary(new_srad)
# 
# srad_flip_fill <- new_srad %>%
#   dplyr::select(-c(doyh)) %>% 
#   pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')
#   
# 
# srad_filled <- ggplot(srad_flip_fill, aes(x=date, y=sw_in_mjm2, color=site)) +
#   geom_line() +
#   ggtitle('Solar rad, fill')
# 
# ggplotly(srad_filled)

@@

```{r}
prec_runoff <- read_csv('C:/Users/sears/Documents/Repos/fourmile/data/prec_runoff.csv')

>>>>

---
title: "Fourmile"
author: "Megan Sears"
date: "2022-10-12"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

@@

#save.image(file = 'spatial.RData')
load('C:/Users/sears/Documents/Repos/fourmile/data/spatial.RData') ## need to update with new locations + fire boundary ***

@@

```{r, echo = F}
niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>%
  filter(!row_number() %in% c(1:10)) %>%
   rename(date = 1,
          prism_mm = 2) %>%
   mutate(date = ymd(date),
          prism_mm = as.numeric(prism_mm))

niwot_snotel <- prec_update %>%
  dplyr::select(c(date, Niwot_Snotel)) %>%
  rename(snotel_mm = Niwot_Snotel) %>%
  mutate(date = dmy(date))

niwot_compare <- full_join(niwot_prism, niwot_snotel, by = 'date')

niwot_compare <- niwot_compare %>%
  mutate(diff_mm = snotel_mm - prism_mm) %>%
  na.omit()

#pbias(niwot_compare$prism_mm, niwot_compare$snotel_mm)

scatter <- ggplot(niwot_compare, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism vs. Niwot SNOTEL precip')

ggplotly(scatter)

rmse(niwot_compare$prism_mm, niwot_compare$snotel_mm)
bias(niwot_compare$snotel_mm, niwot_compare$prism_mm)
summary(niwot_compare$diff_mm)
#filter to look at sept - june
niwot_winter_comp <- niwot_compare %>%
  mutate(month = month(date)) %>%
  filter(!month %in% c(6:8))

summary(niwot_winter_comp)
```

@@

```{r, echo = F}
niwot_prismshift <- niwot_compare %>%
  mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
  mutate(month = month(date)) %>%
  #filter(!month %in% c(6:8)) %>%
  mutate(diff_mm = snotel_mm - prism_mm) %>%
  na.omit()

#NSE(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
#pbias(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)

scatter_shiftday <- ggplot(niwot_prismshift, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day versus Niwot SNOTEL precip')

ggplotly(scatter_shiftday)

rmse(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
bias(niwot_prismshift$snotel_mm, niwot_prismshift$prism_mm)
summary(niwot_prismshift$diff_mm)
```

@@

```{r, echo = F}
niwot_prismfrac <- niwot_compare %>%
  mutate(prism_most_mm = prism_mm * 0.974,
         prism_little_mm = prism_mm * 0.026) %>%
  mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
  mutate(prism_new_mm = prism_most_mm + prism_little_mm,
         diff_prismfrac = snotel_mm - prism_new_mm) %>%
  na.omit()

scatter_frac <- ggplot(niwot_prismfrac, aes(x=snotel_mm, y=prism_new_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 97.4% and Niwot SNOTEL precip')

ggplotly(scatter_frac)

rmse(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
bias(niwot_prismfrac$snotel_mm, niwot_prismfrac$prism_new_mm)
summary(niwot_prismfrac$diff_prismfrac)
NSE(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
pbias(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)

@@

```{r, echo = F}
srad <- solr %>%
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:17)) %>%
  replace(.==-9999, NA) %>%
  rename(date = 1,
         prism_w = 2,
         prism_m = 3,
         prism_e = 4,
         gordon_s = 5,
         gordon_n = 6,
         greenlake = 7) %>%
  mutate(date = dmy(date))

niwot_srad <- Niwo %>%
  mutate(date = ymd(date)) %>%
  rename(niwot_flux_mjm2 = 3) %>%
  dplyr::select(-c(watts_per_m2)) %>%
  replace(. == -9999, NA)

## original
srad_compare <- full_join(srad, niwot_srad, by = 'date') %>%
  dplyr::select(c(date, prism_w, niwot_flux_mjm2)) %>%
  mutate(diff = niwot_flux_mjm2 - prism_w) %>%
  na.omit()

scatter_srad_orig <- ggplot(srad_compare, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism versus Niwot flux tower solar rad')

ggplotly(scatter_srad_orig)

rmse(srad_compare$prism_w, srad_compare$niwot_flux_mjm2)
bias(srad_compare$niwot_flux_mjm2, srad_compare$prism_w)
summary(srad_compare$diff)
```

@@

```{r, echo = F}
## 1 day shift
srad_shift1day <- srad_compare %>%
  mutate_at(c("prism_w"),list(lead), n = 1) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - prism_w)

scatter_srad_1day <- ggplot(srad_shift1day, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_1day)
#scatter_srad_1day

rmse(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
bias(srad_shift1day$niwot_flux_mjm2, srad_shift1day$prism_w)
summary(srad_shift1day$diff)
pbias(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
```

@@

```{r, echo = F}
## shift by percentage
srad_fracshift <- srad_compare %>%
    mutate(prism_most = prism_w * 0.3,
         prism_little = prism_w * 0.7) %>%
  mutate_at(c('prism_most'), list(lead), n = 1) %>%
  mutate(prism_new = prism_most + prism_little,
         diff_prismfrac = niwot_flux_mjm2 - prism_new) %>%
  na.omit()

scatter_srad_frac <- ggplot(srad_fracshift, aes(x=niwot_flux_mjm2, y=prism_new)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 30% vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_frac)
#scatter_srad_frac

rmse(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
bias(srad_fracshift$niwot_flux_mjm2, srad_fracshift$prism_new)
summary(srad_fracshift$diff_prismfrac)
pbias(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
```

@@

```{r, echo = F}
## bring in daymet
# Daily total radiation (MJ/m2/day) can be calculated as follows: ((srad (W/m2) * dayl (s/day)) / l,000,000)
niwot_daymet <- daym %>%
  filter(!row_number() %in% c(1:7)) %>%
  rename(year = 1,
         doy = 2,
         srad_wm2 = 3) %>%
  mutate(date = as.Date(doy),
         month = month(date),
         day = day(date),
         date = make_date(year, month, day),
         daymet_mjm2 = (srad_wm2 * 86400) / 1000000) %>%
  dplyr::select(-c(year, doy, month, day))

niwot_daymet_comp <- full_join(niwot_srad, niwot_daymet, by = 'date') %>%
  dplyr::select(c(date, niwot_flux_mjm2, daymet_mjm2)) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - daymet_mjm2)

scatter_srad_daymet <- ggplot(niwot_daymet_comp, aes(x=niwot_flux_mjm2, y=daymet_mjm2)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Daymet vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_daymet)
#scatter_srad_frac

#summary(srad_fracshift)
rmse(niwot_daymet_comp$daymet_mjm2, niwot_daymet_comp$niwot_flux_mjm2)
bias(niwot_daymet_comp$niwot_flux_mjm2, niwot_daymet_comp$daymet_mjm2)
summary(niwot_daymet_comp$diff)
```

@@

```{r, echo=F}
#bring in srad and niwot_srad
srad_gapfil <- full_join(srad, niwot_srad) %>%
  filter(date > '2009-12-31') %>%
  dplyr::select(-c(prism_w, prism_m, prism_e))

#bat <- read_xlsx('bat_srad.xlsx')
#save.image(file = 'bat.RData')
load('C:/Users/sears/Documents/Repos/fourmile/data/bat.RData')

bet_day <- bat %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(sw_in_avg_wm2= mean(in_swrad_wm2)) %>%
  filter(date > '2009-12-31') %>%
  mutate(bet_sw_in_mjm2 = (sw_in_avg_wm2 * 86400) / 1000000) %>%
  dplyr::select(-c(sw_in_avg_wm2))

srad_gapfil_all <- full_join(srad_gapfil, bet_day, by='date') %>%
  filter(date < '2020-12-31') %>%
  dplyr::select(-c(bet_sw_in_mjm2.x)) %>%
  rename(bet_sw = bet_sw_in_mjm2.y) %>%
  mutate(doyh = yday(date))

summary(srad_gapfil_all)

srad_flip_nofill <- srad_gapfil %>%
  pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')

srad_nofill <- ggplot(srad_flip_nofill, aes(x=date, y=sw_in_mjm2, color=site)) +
  geom_line() +
  ggtitle('Solar rad, no fill')

ggplotly(srad_nofill)

new_srad <- missRanger(srad_gapfil_all, num.trees=100000)
summary(new_srad)

srad_flip_fill <- new_srad %>%
  dplyr::select(-c(doyh)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')

srad_filled <- ggplot(srad_flip_fill, aes(x=date, y=sw_in_mjm2, color=site)) +
  geom_line() +
  ggtitle('Solar rad, fill')

ggplotly(srad_filled)

@@

```{r}
prec_runoff <- read_csv('C:/Users/sears/Documents/Repos/fourmile/data/runoff_prep_precip.csv')
